---
title: "Title of Your Report"
author: "Names of your Group Members"
date: "Due Date"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, echo = FALSE}
library(janitor)
library(tidyverse)

# Load the data dictionary and the raw data and correct the variable names
raw_data <- read_csv("AAqA21FV.csv")
dict <- read_lines("gss_dict.txt", skip = 18) # skip is because of preamble content
# Now we need the labels because these are the actual responses that we need
labels_raw <- read_file("gss_labels.txt")


#### Set-up the dictionary ####
# What we want is a variable name and a variable definition
variable_descriptions <- as_tibble(dict) %>% 
  filter(value!="}") %>% 
  mutate(value = str_replace(value, ".+%[0-9].*f[ ]{2,}", "")) %>% 
  mutate(value = str_remove_all(value, "\"")) %>% 
  rename(variable_description = value) %>% 
  bind_cols(tibble(variable_name = colnames(raw_data)[-1]))
 
# Now we want a variable name and the possible values
labels_raw_tibble <- as_tibble(str_split(labels_raw, ";")[[1]]) %>% 
  filter(row_number()!=1) %>% 
  mutate(value = str_remove(value, "\nlabel define ")) %>% 
  mutate(value = str_replace(value, "[ ]{2,}", "XXX")) %>% 
  mutate(splits = str_split(value, "XXX")) %>% 
  rowwise() %>% 
  mutate(variable_name = splits[1], cases = splits[2]) %>% 
  mutate(cases = str_replace_all(cases, "\n [ ]{2,}", "")) %>%
  select(variable_name, cases) %>% 
  drop_na()

# Now we have the variable name and the different options e.g. age and 0-9, 10-19, etc.
labels_raw_tibble <- labels_raw_tibble %>% 
  mutate(splits = str_split(cases, "[ ]{0,}\"[ ]{0,}"))

# The function sets up the regex (I know, I know, but eh: https://xkcd.com/208/)
add_cw_text <- function(x, y){
  if(!is.na(as.numeric(x))){
    x_new <- paste0(y, "==", x,"~")
  }
  else{
    x_new <- paste0("\"",x,"\",")
  }
  return(x_new)
}

# The function will be in the row, but it'll get the job done
cw_statements <- labels_raw_tibble %>% 
  rowwise() %>% 
  mutate(splits_with_cw_text = list(modify(splits, add_cw_text, y = variable_name))) %>% 
  mutate(cw_statement = paste(splits_with_cw_text, collapse = "")) %>% 
  mutate(cw_statement = paste0("case_when(", cw_statement,"TRUE~\"NA\")")) %>% 
  mutate(cw_statement = str_replace(cw_statement, ",\"\",",",")) %>% 
  select(variable_name, cw_statement)
# So for every variable we now have a case_when() statement that will convert 
# from the number to the actual response.

# Just do some finally cleanup of the regex.
cw_statements <- 
  cw_statements %>% 
  mutate(variable_name = str_remove_all(variable_name, "\\r")) %>% 
  mutate(cw_statement = str_remove_all(cw_statement, "\\r"))


#### Apply that dictionary to the raw data ####
# Pull out a bunch of variables and then apply the case when statement for the categorical variables
gss <- raw_data %>% 
  select(CASEID, 
         agedc, 
         achd_1c, 
         achdmpl, 
         totchdc, 
         acu0c,
         agema1c,
         achb1c,
         rsh_131a,
         arretwk,
         slm_01, 
         sex, 
         brthcan, 
         brthfcan,
         brthmcan,
         brthmacr,
         brthprvc,
         yrarri,
         prv, 
         region, 
         luc_rst, 
         marstat, 
         amb_01, 
         vismin, 
         alndimmg,
         bpr_16, 
         bpr_19,
         ehg3_01b, 
         odr_10, 
         livarr12, 
         dwelc, 
         hsdsizec,
         brthpcan,
         brtpprvc, 
         visminpr,
         rsh_125a, 
         eop_200,
         uhw_16gr,
         lmam_01, 
         acmpryr,
         srh_110,
         srh_115,
         religflg, 
         rlr_110,
         lanhome, 
         lan_01,
         famincg2, 
         ttlincg2, 
         noc1610, 
         cc_20_1,
         cc_30_1,
         ccmoc1c,
         cor_031,
         cor_041,
         cu0rnkc,
         pr_cl,
         chh0014c,
         nochricc,
         grndpa,
         gparliv,
         evermar,
         ma0_220,
         nmarevrc,
         ree_02,
         rsh_131b,
         rto_101,
         rto_110,
         rto_120,
         rtw_300,
         sts_410,
         csp_105,
         csp_110a,
         csp_110b,
         csp_110c,
         csp_110d,
         csp_160,
         fi_110) %>% 
  mutate_at(vars(agedc:fi_110), .funs = funs(ifelse(.>=96, NA, .))) %>% 
  mutate_at(.vars = vars(sex:fi_110),
            .funs = funs(eval(parse(text = cw_statements %>%
                                      filter(variable_name==deparse(substitute(.))) %>%
                                      select(cw_statement) %>%
                                      pull()))))

# Fix the names
gss <- gss %>% 
  clean_names() %>% 
  rename(age = agedc,
         age_first_child = achd_1c,
         age_youngest_child_under_6 = achdmpl,
         total_children = totchdc,
         age_start_relationship = acu0c,
         age_at_first_marriage = agema1c,
         age_at_first_birth = achb1c,
         distance_between_houses = rsh_131a,
         age_youngest_child_returned_work = arretwk,
         feelings_life = slm_01,
         sex = sex,
         place_birth_canada = brthcan,
         place_birth_father = brthfcan,
         place_birth_mother = brthmcan,
         place_birth_macro_region = brthmacr,
         place_birth_province = brthprvc,
         year_arrived_canada = yrarri,
         province = prv,
         region = region,
         pop_center = luc_rst,
         marital_status = marstat,
         aboriginal = amb_01,
         vis_minority = vismin,
         age_immigration = alndimmg,
         landed_immigrant = bpr_16,
         citizenship_status = bpr_19,
         education = ehg3_01b,
         own_rent = odr_10,
         living_arrangement = livarr12,
         hh_type = dwelc,
         hh_size = hsdsizec,
         partner_birth_country = brthpcan,
         partner_birth_province = brtpprvc,
         partner_vis_minority = visminpr,
         partner_sex = rsh_125a,
         partner_education = eop_200,
         average_hours_worked = uhw_16gr,
         worked_last_week = lmam_01,
         partner_main_activity = acmpryr,
         self_rated_health = srh_110,
         self_rated_mental_health = srh_115,
         religion_has_affiliation = religflg,
         regilion_importance = rlr_110,
         language_home = lanhome,
         language_knowledge = lan_01,
         income_family = famincg2,
         income_respondent = ttlincg2,
         occupation = noc1610,
         childcare_regular = cc_20_1,
         childcare_type = cc_30_1,
         childcare_monthly_cost = ccmoc1c,
         ever_fathered_child = cor_031,
         ever_given_birth = cor_041,
         number_of_current_union = cu0rnkc,
         lives_with_partner = pr_cl,
         children_in_household = chh0014c,
         number_total_children_intention = nochricc,
         has_grandchildren = grndpa,
         grandparents_still_living = gparliv,
         ever_married = evermar,
         current_marriage_is_first = ma0_220,
         number_marriages = nmarevrc,
         religion_participation = ree_02,
         partner_location_residence = rsh_131b,
         full_part_time_work = rto_101,
         time_off_work_birth = rto_110,
         reason_no_time_off_birth = rto_120,
         returned_same_job = rtw_300,
         satisfied_time_children = sts_410,
         provide_or_receive_fin_supp = csp_105,
         fin_supp_child_supp = csp_110a,
         fin_supp_child_exp = csp_110b,
         fin_supp_lump = csp_110c,
         fin_supp_other = csp_110d,
         fin_supp_agreement = csp_160,
         future_children_intention = fi_110) 

#### Clean up ####
gss <- gss %>% 
  mutate_at(vars(age:future_children_intention), 
            .funs = funs(ifelse(.=="Valid skip"|.=="Refusal"|.=="Not stated", "NA", .))) 

gss <- gss %>% 
  mutate(is_male = ifelse(sex=="Male", 1, 0)) 

gss <- gss %>% 
  mutate_at(vars(fin_supp_child_supp:fin_supp_other), .funs = funs(case_when(
    .=="Yes"~1,
    .=="No"~0,
    .=="NA"~as.numeric(NA)
  )))

main_act <- raw_data %>% 
  mutate(main_activity = case_when(
    mpl_105a=="Yes"~ "Working at a paid job/business",
    mpl_105b=="Yes" ~ "Looking for paid work",
    mpl_105c=="Yes" ~ "Going to school",
    mpl_105d=="Yes" ~ "Caring for children",
    mpl_105e=="Yes" ~ "Household work", 
    mpl_105i=="Yes" ~ "Other", 
    TRUE~ "NA")) %>% 
  select(main_activity) %>% 
  pull()

age_diff <- raw_data %>% 
  select(marstat, aprcu0c, adfgrma0) %>% 
  mutate_at(.vars = vars(aprcu0c:adfgrma0),
            .funs = funs(eval(parse(text = cw_statements %>%
                                      filter(variable_name==deparse(substitute(.))) %>%
                                      select(cw_statement) %>%
                                      pull())))) %>% 
  mutate(age_diff = ifelse(marstat=="Living common-law", aprcu0c, adfgrma0)) %>% 
  mutate_at(vars(age_diff), .funs = funs(ifelse(.=="Valid skip"|.=="Refusal"|.=="Not stated", "NA", .))) %>% 
  select(age_diff) %>% 
  pull()

gss <- gss %>% mutate(main_activity = main_act, age_diff = age_diff)

# Change some from strings into numbers
gss <- gss %>% 
  rowwise() %>% 
  mutate(hh_size = str_remove(string = hh_size, pattern = "\\ .*")) %>% 
  mutate(hh_size = case_when(
    hh_size=="One" ~ 1,
    hh_size=="Two" ~ 2,
    hh_size=="Three" ~ 3,
    hh_size=="Four" ~ 4,
    hh_size=="Five" ~ 5,
    hh_size=="Six" ~ 6
  )) 

gss <- gss %>% 
  rowwise() %>% 
  mutate(number_marriages = str_remove(string = number_marriages, pattern = "\\ .*")) %>% 
  mutate(number_marriages = case_when(
    number_marriages=="No" ~ 0,
    number_marriages=="One" ~ 1,
    number_marriages=="Two" ~ 2,
    number_marriages=="Three" ~ 3,
    number_marriages=="Four" ~ 4
  )) 

gss <- gss %>% 
  rowwise() %>% 
  mutate(number_total_children_known = ifelse(number_total_children_intention=="Don't know"|number_total_children_intention=="NA", 0, 1)) %>% 
  mutate(number_total_children_intention = str_remove(string = number_total_children_intention, pattern = "\\ .*")) %>% 
  mutate(number_total_children_intention = case_when(
    number_total_children_intention=="None" ~ 0,
    number_total_children_intention=="One" ~ 1,
    number_total_children_intention=="Two" ~ 2,
    number_total_children_intention=="Three" ~ 3,
    number_total_children_intention=="Four" ~ 4,
    number_total_children_intention=="Don't" ~ as.numeric(NA)
  )) 

write_csv(gss, "gss.csv")
```

## Title of your Report

# Yunrui Feng, Kekun Han, Yichen Liang 
# Date

## Abstract
According to a five-year study from Edith Cowan University in Australia, families with more children have a higher level of life satisfaction rather than feeling overwhelmed by the big family size. The study interviewed hundreds of parents from families of different structures. They found that larger families benefit from increased support and therefore have the highest overall life satisfaction. Our aim in this report is to investigate whether we can observe similar associations between life satisfaction and household size. To test the hypothesis that there is a positive correlation between household size and life satisfaction, we make a choice of using a linear regression model to analyze their quantitative relationship as they are both continuous variables. The dependent variable is life satisfaction and the independent variable is household size. The analysis does not show a statistically significant result on the correlation between life satisfaction and household size, but it does show that household size larger than one has a higher level of life satisfaction than a single person. We conclude that household size larger than one has a higher level of life satisfaction.


## Introduction
Feelings of life, in other words, life satisfaction is an overall assessment of feelings and attitudes about one's life as a whole. Numerous studies show that life satisfaction has a positive impact on our health, and it is one of the major indicators of well-being. For instance, researchers at Chapman University published a research on life satisfaction and its association with reducing risk of mortality(Morgan). Regarding the importance of improving life satisfaction, we are interested in the factors associated with the changes in the level of life satisfaction. We use the data from the Public Use Microdata File (PUMF) of the 2017 General Social Survey (GSS) on the Family. We use the ‘gss_cleaning’ code provided to clean up the data and put it into a tidy format. Since we are investigating the relationship between life satisfaction and household size, we extract the columns containing the corresponding data and put them in a data frame for analysis. In the report, we create plots and investigate mean life satisfaction scores for each size of the household, confidence intervals,  interquartile range of the residuals, estimated coefficients along with its standard error and significance test, and other model statistics. We discuss the strengths and weaknesses of the data and come to a conclusion. 




## Data
##### The data used for analysis is from the 2017 General Social Survey. The target population of this survey includes "all non-institutionalized persons 15 years of age and older living in the ten provinces of Canada". The frame population is created using the lists of telephone numbers in use available to Statistics Canada and The Address Register. And the survey uses stratified random sampling to collect the data. There are two columns of data that are selected, one measures household size ("household_size") and the other one measures the household's feelings about life ("feelings_life). The variable "household_size" has only integer values from 1 to 6, the variable "feelings_life" also only has integer values but it is in a scale of 1 to 10.

##### There are many studies which study the factors that affect an individual's life satisfaction. And some of them conclude that people who live in larger families have higher overall life satisfaction. So it will be interesting to use some data to test this conclusion and try to make some new inferences. Moreover, the dataset has very large number of observations and there is no na value within those observation. So the selected data can be modlled and analysed without too much "cleaning" work. 

##### Furthermore, the selected data also has its drawbacks. For example, this dataset only contains discrete variables and the variable "household_size" will have to be treated as a categorical variable. So it will be harder to fit a model into the data nicely, thus negatively affects the model's ability to predict the dependent variable "feelings_life".

## Model
```{r, echo = FALSE}
happyness<- tibble(gss$hh_size,gss$feelings_life)
happyness <- happyness[complete.cases(happyness),]
happyness <-
  happyness%>%
  rename(household_size = `gss$hh_size`,
         feelings_life = `gss$feelings_life`)
plot(as.factor(happyness$household_size),happyness$feelings_life, xlab = "Household Size", ylab = "Happyness Score", main = "Distribution of life satisfaction scores in each household size")
my_LRmodel <- lm(happyness$feelings_life ~ as.factor(happyness$household_size))
summary(my_LRmodel)
r <- rstudent(my_LRmodel)
plot(happyness$household_size, r, xlab = "Household Size", ylab = "Standardized Residuals",
     main = "Standardized Residuals vs. Household Size")
#她说要draw on relevant literature, 我刚好找到一篇关于这个的“Study finds larger families have highest overall life satisfaction”
```
 
##### The model used here is a mutiple linear regression model: feelings_life = 7.69396 + 0.61378x1 + 0.40607x2 + 0.54916x3 + 0.52602x4 + 0.67527x5. 
##### Note: x1, x2, x3, x4 and x5 are all dummy variables, x1 is 1 only when household size is 2; x2 is 1 only when household size is 3; x3 is 1 only when household size is 4; x4 is 1 only when household size is 5; x4 is 1 only when household size is 5; x5 is 1 only when household size is 6.
##### A linear regression model is used here rather than a logistic regression model because the dependent variable "feelings_life" is not binary. And a linear regression model is good to compare the mean differences of life satisfaction scores between each size of the household. It is useful to find out whether there is a general pattern between household size and life satisfaction. Moreover, population correction is not used because the sample size is less than 5% of the total population.
##### For diagnosis checking, because all independent variables are dummy variables, the linearity assumption is certainly satisfied. The residual plot does not show any patterns or specific clusters of data points thus both the homoscedasticity and normality assumptions are met. Therefore, this should be a valid model to analyze further.


## Results

```{r, echo = FALSE}
summary(my_LRmodel)
```

#### This summary contains the results of the estimated model includes the interquartile range of the residuals, the estimated coefficients along with its standard error and significance test, and other model statistics such as R-squared. This summary will be called table 1.

```{r, echo = FALSE}
avg_hz2 <- as.numeric(my_LRmodel$coefficients[2]) + as.numeric(my_LRmodel$coefficients[1]) 
avg_hz3 <- as.numeric(my_LRmodel$coefficients[3]) + as.numeric(my_LRmodel$coefficients[1]) 
avg_hz4 <- as.numeric(my_LRmodel$coefficients[4]) + as.numeric(my_LRmodel$coefficients[1]) 
avg_hz5 <- as.numeric(my_LRmodel$coefficients[5]) + as.numeric(my_LRmodel$coefficients[1]) 
avg_hz6 <- as.numeric(my_LRmodel$coefficients[6]) + as.numeric(my_LRmodel$coefficients[1])

upper_confidence_interval_hz1 <- as.numeric(my_LRmodel$coefficients[1]) + 1.96*0.02149
upper_confidence_interval_hz2 <- avg_hz2 + 1.96*0.02840  
upper_confidence_interval_hz3 <- avg_hz3 + 1.96*0.03726 
upper_confidence_interval_hz4 <- avg_hz4 + 1.96*0.03782 
upper_confidence_interval_hz5 <- avg_hz5 + 1.96*0.05591 
upper_confidence_interval_hz6 <- avg_hz6 + 1.96*0.08504 

lower_confidence_interval_hz1 <- as.numeric(my_LRmodel$coefficients[1]) - 1.96*0.02149
lower_confidence_interval_hz2 <- avg_hz2 - 1.96*0.02840 
lower_confidence_interval_hz3 <- avg_hz3 - 1.96*0.03726 
lower_confidence_interval_hz4 <- avg_hz4 - 1.96*0.03782 
lower_confidence_interval_hz5 <- avg_hz5 - 1.96*0.05591 
lower_confidence_interval_hz6 <- avg_hz6 - 1.96*0.08504

avg_hz <- matrix(c(as.numeric(my_LRmodel$coefficients[1]),lower_confidence_interval_hz1, upper_confidence_interval_hz1, avg_hz2, lower_confidence_interval_hz2,upper_confidence_interval_hz2,  avg_hz3, lower_confidence_interval_hz3, upper_confidence_interval_hz3,  avg_hz4, lower_confidence_interval_hz4, upper_confidence_interval_hz4, avg_hz5, lower_confidence_interval_hz5, upper_confidence_interval_hz5, avg_hz6, lower_confidence_interval_hz6, upper_confidence_interval_hz6), ncol = 3, byrow = TRUE)
colnames(avg_hz) <- c("AVG_Feelings_Life", "95% CI (Lower Bound)", "95% CL (Upper Bound)")
rownames(avg_hz) <- c("Household_size_1","Household_size_2","Household_size_3","Household_size_4","Household_size_5","Household_size_6")
avg_hz <- as.table(avg_hz)
avg_hz
```

#### This table contains the mean life satisfaction scores for each size of the household and their corresponding 95% confidence intervals. And it will be called table 2.


```{r, echo = FALSE}
ggplot(happyness, aes(x = household_size,y=feelings_life))+geom_point()+geom_smooth(method=lm)+ labs(title="Scatterplot of household size versus feeling about life") +  xlim(1,6) + ylim(0,10)
```

#### This graph shows the fitted value visually by adding the model prediction (blue line) to the scatter plot of the original dataset. And it will be called figure 3.


## Discussion
After detailed analyzation, model construction and graph plotting, there are several meaningful and conclusive statements to be made. 
Our original null hypothesis is that there is a positive correlation between household size and life satisfaction, after we have constructed a simple linear regression model that fits the data, we observe that the correlation between house size and the level of life satisfaction do not have a strong correlation. Referring back to summary table 1, we can see that the values standard error of regression for each house size are relatively low and close to 0. As we know, the standard error of the regression provides the absolute measure of the typical distance that the data points fall from the regression line, since we have low values for this statistic, it suggests that the data points and the fitted values are distanced very closely, suggesting that the model is a very good fit for the regression model. However, when we analyze the t-value for each of the household sizes, we can see that the actual correlation is very weak. The t-value measures the size of the difference comparing to the variation in the sample data,  and when the value is large, it suggests that there is significant difference, which gives us stronger difference against the null hypothesis. As we can see from table 1, the t-value for each of the household sizes range from approximtely 8 to approximately 21, which are all relatively far away from 0, suggesting that the evidence against the null hypothesis of a positive relationship existing between the household size and the level of life satisfaction is very strong, giving us evidence to reject the original hypothesis. 
On the other hand, when we interpret the p-values, we can draw similar conclusions. Referring back to table again, we can see that the p-values are all relatively small and are all less than 0.05. In addition, when we look at figure 3, we can see that the slope of correlation between household size and feelings about life is relatively flat, although it is a positive relationship, we can draw the conclusion that the correlation is not strong. 
In conclusion, statistically speaking, there is no significant result on the correlation between life satisfaction and household size. 

When we directly compare the distribution of the level of life satisfaction between different household sizes, we can draw some different conclusions. When we look at the Distribution of life satisfaction scores in each household size, we can see that average level of life satisfaction increases as household size increases, suggesting that there as the number of people in a family increases, the level of happiness do increase. Furthermore,  when we analyze the 95% confidence interval of the data,  we can see that a household of 6 has the highest score while a household of 1 has the lowest score.  In addition, we can see that lower bound and higher bound of the confidence interval increases as the number of people in a household increases, suggesting once again that the level of life satisfaction is correlated with the household size. 

In general conclusion, we did not find direct statistical evidence to support the hypothesis that there is a strong positivef correlation between the level of life satisfaction and the household sizes. However, comparatively speaking, there is enough data to suggest that a family is happier with more people. This result tells us that a happy and satisfactory life is always supported by a well-functioning family, and that the lack of interaction between family members could affect the level of life satisfaction for any individual in life. 

# Weaknesses

One of the most significant weakness of the study is that the data collected could not fully reflect the level of life satisfaction. As we know, the data is collected using surveys, and sometimes survey questions are too inflexible that it could not fully represent and demonstrate the intentions behind the questionnaire. Survey questions could also be confusing to respondents, therefore the results recorded could reflect a biased response. In addition,  we can see that the statistical results we obtained from the linear regression model is inconsistent with the results we got when we directly compare the data, this is a major problem and weakness. This could be result that there are many outliers in the dataset, as we can see from the Distribution of life satisfaction scores in each household size, there are many data entries that are well below the first interquartile range, and this could potentially affect the statistical analysis in many ways. For improvements, we could attempt to extract all the outliers from the dataset and then study the specific correlation with only the relevant data scores. In addition, we all know that the level  of life satisfaction is also deeply connnected with other factors such as income, therefore a simple linear regression model with only one variable of household size might not fully reflect the level of happiness in an individual's life. 

# Next Steps

As discussed in the previous section, our next step could be to analyze the correlation between level of life satisfaction and multiple variables (such as household size, level of income, educational level, etc). We would do so by constructing a multiple regression model where we could also study the correlation between different variables. One of the reasons that the statistical result we obtained in this study do not match with the observation might be that there is other correlations that we did not account for in a simple linear regression model, therefore it would be beneficial to study upon that in the subsequent work to be done after this report. 




## References
https://www150.statcan.gc.ca/n1/daily-quotidien/170913/t001a-eng.htm

https://sda-artsci-utoronto-ca.myaccess.library.utoronto.ca/sdaweb/dli2/gss/gss31/gss31/more_doc/GSS31_User_Guide.pdf


Morgan, M. (2017, March 03). Study finds larger families have highest overall life satisfaction. Retrieved October 18, 2020, from https://www.reviewjournal.com/life/study-finds-larger-families-have-highest-overall-life-satisfaction/